% rubber: bibtex.crossrefs 10000
\documentclass[a4paper,12pt,oneside,minionpro,dottedtoc]{thesis}

\usepackage[a4paper,left=2cm,right=2cm,top=2cm,bottom=3cm]{geometry}

\title{Multi-version Execution for Increasing the Reliability and Availability of Updated Software}
\author{Petr Hosek}
\date{August 2015}
\university{Imperial College London}
\department{Department of Computing}
\supervisor{Dr Cristian Cadar}
\dedication{To everyone who believed in me when others did not.} 
%\dedication{I would like to dedicate this thesis to everyone who believed in me when others did not.
%  Special mention goes to Elena, Cristian, Susan and Google. You have my undying gratitude.} 
\declaration{This thesis is a presentation of my own original research and all the sources
have been quoted and acknowledged by means of complete references.}
\submitdate{30 August 2015}

\lstset{
  language=C,
  captionpos=b,
  basicstyle=\small\ttfamily,
  commentstyle=\textit,
  keywordstyle=\bfseries,
  numbers=left,
  numberstyle=\small,
  numbersep=5pt,
  escapeinside={/*@}{@*/},
  tabsize=2,
  breaklines=true,
  extendedchars=true,
  columns=fullflexible,
  xleftmargin=16pt,
  showspaces=false,
  showtabs=false,
  showstringspaces=false,
  moredelim=**[is][\bfseries]{@}{@},
}

\lstdefinelanguage
   [x64]{Assembler}
   [x86masm]{Assembler}
   {morekeywords={CDQE,CQO,CMPSQ,CMPXCHG16B,JRCXZ,LODSQ,MOVSXD,%
                  POPFQ,PUSHFQ,SCASQ,STOSQ,IRETQ,RDTSCP,SWAPGS,%
                  SYSCALL,SYSRET,SYSENTER,SYSEXIT,%
                  rax,rdx,rcx,rbx,rsi,rdi,rsp,rbp,%
                  r8,r8d,r8w,r8b,r9,r9d,r9w,r9b}}

\lstdefinelanguage
  [bpf]{Assembler}
  {morekeywords={ldb,ldh,ld,ldi,ldx,ldxi,ldxb,st,stx,jmp,ja,jeq,jneq,%
                 jne,jlt,jle,jgt,jge,jset,add,sub,mul,div,mod,neg,and,xor,or,%
                 lsh,rsh,ret,tax,txa},%
   alsoletter=.,alsodigit=?,%
   sensitive=false,%
   morestring=[b]",%
   morestring=[b]',%
   morecomment=[l];%
   }[keywords,comments,strings]

\lstdefinelanguage{diff}{
   moredelim=[s][\bfseries]{@@}{@@},%
   moredelim=*[l]-,%
   moredelim=*[l]+,%
   moredelim=*[l]{---},%
   moredelim=*[l]{+++},%
}

%\usemintedstyle{bw}

\usepackage[binary-units,group-separator={,}]{siunitx}
\sisetup{per-mode=symbol}
\DeclareSIUnit[number-unit-product=\,]{\loc}{LOC}
\DeclareSIUnit[number-unit-product=\,]{\eloc}{ELOC}
\DeclareSIUnit[number-unit-product=\,]{\reqs}{reqs}

\usepackage{xspace}
\usepackage{url}
\usepackage[normalem]{ulem}
\usepackage{multirow}
\usepackage{paralist}
\setlength{\parskip}{0cm}
\setlength{\parindent}{1em}

\usepackage[backend=biber,style=alphabetic,mincrossrefs=10000]{biblatex}
\addbibresource{bibliography/macros.bib}
\addbibresource{bibliography/crossrefs.bib}
\addbibresource{bibliography/refs.bib}

\newcommand*\circl[1]{%
  \begin{tikzpicture}[baseline=(C.base)]
    \node[draw,circle,inner sep=0.25pt](C) {\small\textsf{#1}};
  \end{tikzpicture}}

\newcommand{\boldtext}[1]{\vspace{0.05in}\noindent\textbf{#1}}
\newcommand{\textstt}[1]{{\small \texttt{#1}}\xspace}
%\newcommand{\textssc}[1]{{\small \textsc{#1}}\xspace}
\newcommand{\textssc}[1]{\textls[80]{\scshape\MakeTextLowercase{#1}}}

\newcommand{\mx}[0]{\textsc{Mx}\xspace}
\newcommand{\mxe}[0]{\textsc{MxE}\xspace}
\newcommand{\sea}[0]{\textsc{Sea}\xspace}
\newcommand{\rem}[0]{\textsc{Rem}\xspace}
\newcommand{\mxm}[0]{\textsc{Mxm}\xspace}
\newcommand{\mv}[0]{\textsc{MV}\xspace}

\newcommand{\ptrace}[0]{\textsf{ptrace}\xspace}
\newcommand{\apachebench}[0]{\textstt{ApacheBench}\xspace}
\newcommand{\redisbenchmark}[0]{\textstt{redis-benchmark}\xspace}
\newcommand{\beanstalkdbenchmark}[0]{\textstt{beanstalkd-benchmark}\xspace}
\newcommand{\wrk}[0]{\textstt{wrk}\xspace}
\newcommand{\memslap}[0]{\textstt{memslap}\xspace}
\newcommand{\libmemcached}[0]{\emph{libMemcached}\xspace}
\newcommand{\httpload}[0]{\textstt{http\_load}\xspace}
\newcommand{\vsftpd}[0]{\emph{vsftpd}\xspace}
\newcommand{\vim}[0]{\emph{Vim}\xspace}
\newcommand{\gchrome}[0]{\emph{Google Chrome}\xspace}
\newcommand{\chrome}[0]{\emph{Chrome}\xspace}
\newcommand{\gnu}[0]{\textit{GNU}\xspace}
\newcommand{\coreutils}[0]{\emph{Coreutils}\xspace}

\newcommand{\binutils}[0]{\textit{Binutils}\xspace}
\newcommand{\mdsum}[0]{\textstt{md5sum}\xspace}
\newcommand{\shasum}[0]{\textstt{sha1sum}\xspace}
\newcommand{\mkdir}[0]{\textstt{mkdir}\xspace}
\newcommand{\mkfifo}[0]{\textstt{mkfifo}\xspace}
\newcommand{\mknod}[0]{\textstt{mknod}\xspace}
\newcommand{\cut}[0]{\textstt{cut}\xspace}

\newcommand{\todo}[1]{\noindent{\color{red}[{\bf todo:} #1]}}
\newcommand{\sidenote}[1]{{\marginpar{\bf\color{red}[#1]}}}
\newcommand{\gap}[2]{\uline{#1} \noindent{\color{red}[#2]}}

\newenvironment{structure}{\color{blue}
\begin{list}{\labelitemi}{\leftmargin=1em}{%
\setlength{\topsep}{0pt}\setlength{\itemsep}{0pt}\setlength{\parsep}{0pt}}}
{\end{list}}

\newenvironment{structure*}{\color{blue}
\begin{enumerate}\setlength{\topsep}{0pt}\setlength{\itemsep}{0pt}\setlength{\parsep}{0pt}}
{\end{enumerate}}

\newcommand{\ie}{\emph{i.e.}\ }
\newcommand{\eg}{\emph{e.g.},\ }
\newcommand{\etal}{\emph{et al.}\ }
\newcommand{\vs}{\emph{vs.}\ }
\newcommand{\etc}{\emph{etc.}\ }
\newcommand{\resp}{\emph{resp.}\ }

\newcommand{\sref}[1]{\S\ref{#1}}

%%%

\newcounter{enumcount}
\newenvironment{enumnoind}
 {\begin{list}{(\arabic{enumcount})~~}{\usecounter{enumcount} 
  \labelsep=0em \labelwidth=0em \leftmargin=0em \itemindent=0em}}
 {\end{list}}

\definecolor{lightGrey}{rgb}{0.8, 0.8, 0.8}
\newcommand{\Hilight}{\makebox[0pt][l]{\color{lightGrey}\rule[-0.45em]{\linewidth}{1.5em}}}
\newcommand{\topic}{}

\hyphenation{ex-e-cu-tion}
\hyphenation{de-ter-min-is-tic}
\hyphenation{de-ter-min-is-tic-ally}
\hyphenation{non-de-ter-min-is-tic}
\hyphenation{non-de-ter-min-is-tic-ally}

\newcommand{\stt}[1]{\texttt{\small #1}}
\newcommand{\heading}[1]{\vspace{0.1in} \noindent \textit{#1}}

\newcommand{\hl}[1]{\textit{\textbf{#1}}}
%\newcommand{\hl}[1]{#1}

\newcommand{\diff}{\stt{diff}\xspace}
\newcommand{\gcov}{\stt{gcov}\xspace}
\newcommand{\eloc}{\xspace{\small ELOC}\xspace}
\newcommand{\tloc}{\xspace{\small TLOC}\xspace}

\newcommand{\diffutils}{Diffutils\xspace}
\newcommand{\findutils}{Findutils\xspace}
\newcommand{\zeromq}{{\O}MQ\xspace}
\newcommand{\git}{\textit{Git}\xspace}

\newcommand{\true}{\stt{true}\xspace}
\newcommand{\false}{\stt{false}\xspace}

\newcommand{\nx}[0]{\textsc{Varan}\xspace}
\newcommand{\vx}[0]{\textsc{Vx}\xspace}
\newcommand{\varan}[0]{\textsc{Varan}\xspace}
\newcommand{\nxlib}[0]{\textsc{VaranLib}\xspace}

\newcommand{\linux}{Linux\xspace}
\newcommand{\docker}{Docker\xspace}
\newcommand{\vagrant}{Vagrant\xspace}
\newcommand{\packer}{Packer\xspace}
\newcommand{\fabric}{Fabric\xspace}
\newcommand{\lxc}{LXC\xspace}
\newcommand{\github}{GitHub\xspace}
\newcommand{\cloudstack}{CloudStack\xspace}
\newcommand{\xen}{Xen\xspace}
\newcommand{\kvm}{KVM\xspace}

\newcommand{\seccompbpf}{\textsf{seccomp/bpf}\xspace}
\newcommand{\strace}{\textsf{strace}\xspace}
\newcommand{\perf}{\textsf{perf}\xspace}
\newcommand{\mpstat}{\textsf{mpstat}\xspace}

\newcommand{\apache}[0]{\textit{Apache httpd}\xspace}
\newcommand{\httpd}[0]{\textit{Apache httpd}\xspace}
\newcommand{\beanstalkd}[0]{\textit{Beanstalkd}\xspace}
\newcommand{\lighttpdgen}[0]{\textit{Lighttpd}\xspace}
\newcommand{\lighttpd}[0]{\textit{Lighttpd~1.4}\xspace}
\newcommand{\lighttpdtwo}[0]{\textit{Lighttpd~2.0}\xspace}
\newcommand{\memcached}[0]{\textit{Memcached}\xspace}
\newcommand{\nginx}[0]{\textit{Nginx}\xspace}
\newcommand{\redis}[0]{\textit{Redis}\xspace}
\newcommand{\speccpu}[0]{\textit{SPEC CPU}\xspace}
\newcommand{\speczerozero}[0]{\textit{SPEC CPU2000}\xspace}
\newcommand{\speczerosix}[0]{\textit{SPEC CPU2006}\xspace}
\newcommand{\thttpd}[0]{\textit{thttpd}\xspace}
\newcommand{\gdb}[0]{\textit{GDB}\xspace}

\newcommand{\orchestra}[0]{\textit{Orchestra}\xspace}
\newcommand{\scribe}[0]{\textit{Scribe}\xspace}
\newcommand{\strata}[0]{\textit{Strata}\xspace}
\newcommand{\tachyon}[0]{\textit{Tachyon}\xspace}

\usepackage{hyperref}
\hypersetup{
  unicode=true,
  bookmarksopen=true,
  bookmarksnumbered=true,
  linktoc=page,
  colorlinks=true,
  linkcolor=halfgray,
  citecolor=halfgray,
  urlcolor=halfgray
}

\input{evolution/results}
\input{safe-updates/results}
\input{efficient-execution/results}

\pretolerance=1000
\clubpenalty=10000 
\widowpenalty=10000

\begin{document}
\let\question\savedquestion

\maketitle

\cleardoublepage
\thispagestyle{plain}

\section*{Declaration of Originality}
This thesis is a presentation of my own original research and all the sources
have been quoted and acknowledged by means of complete references.

\section*{Copyright Declaration}
The copyright of this thesis rests with the author and is made available under a Creative Commons
Attribution Non-Commercial No Derivatives licence. Researchers are free to copy, distribute or
transmit the thesis on the condition that they attribute it, that they do not use it for commercial
purposes and that they do not alter, transform or build upon it. For any reuse or redistribution,
researchers must make clear to others the licence terms of this work.

\vfill
\section*{Colophon}

This document was typeset using Lua\LaTeX, using a custom style based on
\textsf{classicthesis}.\footnote{\url{https://code.google.com/p/classicthesis/}}
The bibliography was processed by
Biblatex\footnote{\url{http://www.ctan.org/pkg/biblatex}} and
Biber.\footnote{\url{http://biblatex-biber.sourceforge.net/}}

The default text font is Minion Pro, while the \textsf{sans-serif} text is
typeset in \textsf{Myriad Pro}; \texttt{monospaced} text uses \texttt{Bitstream
Vera Mono}.

\preface
\begin{abstract}
Software updates are an integral part of the software development and
maintenance process, but unfortunately they present a high risk, as new
releases often introduce new bugs and security vulnerabilities; as a
consequence, many users refuse to upgrade their software, relying instead on
outdated versions, which often leave them exposed to known software bugs and
security vulnerabilities.

In this thesis we propose a novel \emph{multi-version execution} approach, a variant
of N-version execution, for improving the software update process. Whenever a
new program update becomes available, instead of upgrading the software to the
newest version, we run the new version in parallel with the old one, and
carefully synchronise their execution to create a more reliable multi-version
application.

We propose two different schemes for implementing the multi-version execution
technique—via failure recovery and via transparent failover—and we describe two
possible designs for implementing these schemes: \mx, focused on recovering
from crashes caused by the faulty software updates; and \varan, focused on
running a large number of versions in parallel with a minimal performance
overhead.

\mx uses static binary analysis, system call interposition, lightweight
checkpointing and runtime state manipulation to implement a novel fault
recovery mechanism, which enables the recovery of the crashing version using
the code of the other, non-crashing version. We have shown how \mx can be
applied successfully to recover from real crashes in several real
applications.

\varan combines selective binary rewriting with high-performance event
streaming to significantly reduce performance overhead, without sacrificing
the size of the trusted computing base, nor flexibility or ease of debugging.
Our experimental evaluation has demonstrated that \varan can run C10k network
servers with low performance overhead, and can be used in various scenarios
such as transparent failover and live sanitisation.
\end{abstract}

\makededication

\begin{acknowledgments}
\noindent
First and foremost, I would like to thank my supervisor Dr Cristian Cadar for
all his support and advice throughout the past four years. He was an excellent
advisor and I have thoroughly enjoyed our collaboration, and I have learned
more than I could have ever imagined.

\noindent
Second, I would like to thank Dr Susan Eisenbach for sharing her wisdom and
experience. She was a great mentor, especially in the beginning of my studies,
and she made me feel like a true part of our department.

\noindent
I would not be able to pursue my studies without Google's financial support; I
was a recipient of the Google Europe Fellowship in Software Engineering, and
this research was supported by this Google Fellowship. I cannot thank them
enough for their generosity.

\noindent
Finally, I would like to thank Elena for all her love and support throughout my
studies which made the last few years so much more enjoyable.
\end{acknowledgments}

\tableofcontents
\listoftables
\listoffigures
\lstlistoflistings

\cleardoublepage

\body

\input{overview/overview}
\input{multi-version/multi-version}
\input{evolution/evolution}

\input{safe-updates/safe-updates}
\input{efficient-execution/efficient-execution}

\input{related/related}
\input{applications/applications}
\input{conclusion/conclusion}
\addcontentsline{toc}{chapter}{References}
{
  \printbibliography
}

\appendix

\end{document}
