\chapter{Efficient Multi-version Execution}

Recent years have seen a growing interest in using diversity as a way
to increase the reliability and security of software systems.  One
form of software diversity that has attracted significant interest
from the research community is the idea of running multiple
diversified versions of a program in parallel in order to survive bugs
and detect security attacks.  In essence, diversity can offer
probabilistic guarantees that at least one variant survives a bug, or
that a security attack will be flagged by divergent behaviour across
variants.

On the reliability side, which forms the main focus of this paper,
these diversified versions are either automatically-generated
variants, multiple versions of the same application, or different
programs implementing the same interface.  For example, one may run in
parallel multiple variants that employ complementary thread schedules
to survive concurrency errors~\cite{compl-schedules11}, multiple
versions of the same software to survive update bugs~\cite{mx}, or
multiple web browsers to benefit from the fact that many errors do not
affect all browser implementations~\cite{cocktail}.  In this paper, we
show that running multiple versions in parallel can be used in other
reliability scenarios, such as running expensive error detectors
(``sanitizers'') during deployment.

On the security side, these diversified variants are constructed in
such a way as to reduce the probability of an attack succeeding in all
of
them~\cite{cox2006,orchestra09,diehard06,tightlip,capizzi08,devries10,cocktail,trachsel10}.
For example, one may generate versions with a different arrangement of
memory blocks in the address space~\cite{diehard06}, or with stacks
growing in opposite directions~\cite{orchestra09}, to prevent attacks
whose success depends on the memory layout.

To enable these scenarios, a monitor process coordinates the parallel
execution of these variants\footnote{The terms \textit{version} and
  \textit{variant} are used interchangeably.} and synchronises their
execution, making them appear as a single application to any outside
entities.  While synchronisation can be performed at different levels,
the most common approach is to do it at the level of system calls, for
two main reasons: first, many existing diversification
transformations, such as the ones discussed above, 
% address-space layout randomisation~\cite{diehard06} and
% instruction-set randomisation~\cite{instr-set-rand03}
do not change the sequence of system calls (the program's
\textit{external behaviour}), and the ordering is often preserved even
across different software versions~\cite{mx}.  Second, system calls
are the main way in which the application communicates with the
outside environment, and therefore
%% the ultimate target of attackers.  Finally, as the main
%% communication mechanism between applications and the environment,
%% system calls
must be virtualised in order to enable the multiple versions to act as
one to the outside world.

The main challenge in implementing an NVX monitor at the system call
level is the trade-off between performance, security, flexibility and
ease of debugging.  Many
implementations~\cite{orchestra09,mx,process-replicae07} use the
\lstinline`ptrace` mechanism offered by most UNIX-based operating
systems.  While easy-to-use and not requiring kernel modifications,
\lstinline`ptrace` is slow, and these systems see performance
degradations of up to two orders of magnitude.  A much faster approach
is to implement the monitor in kernel space~\cite{cox2006}, but this
requires kernel patches and/or new kernel modules, and the monitor
must be run in privileged mode.  Furthermore, none of these approaches
scales well with the number of variants (as the monitor is both a
communication and synchronisation bottleneck), none are debug-friendly
(\lstinline`ptrace` disallows the use of \gdb, while kernel debugging
has its well-known set of limitations) and none of them have been
designed to be flexible with respect to small variations in system
call sequences (which can occur for certain diversification
transformations and across program versions).

In this chapter, we propose \varan,\footnote{\varan's name comes from
  the scientific name \emph{Varanus}, commonly known as the
  \emph{monitor} lizard. Varan is also a name of the Kaiju monster
  that first appeared in the 1958 movie \emph{Varan the
    Unbelievable}.} a novel architecture for implementing NVX
monitors.  \varan monitors operate at the system call level, run in
user space (and therefore in unprivileged mode), introduce a small
performance overhead for popular C10k network servers and a negligible
overhead for CPU-bound applications, scale well with the number of
versions, and provide a flexible mechanism for handling small
divergences in the system call sequences issued across versions.

%\todo{Explicit list of contributions?}

%The rest of this paper is structured as follows.
%Section~\ref{sec:overview} gives a high-level overview of our
%approach, while Section~\ref{sec:prototype} presents our prototype
%implementation in detail.  Then, Section~\ref{sec:evaluation}
%evaluates our prototype on a set of micro- and macro-benchmarks,
%Section~\ref{sec:applications} shows the applicability to different
%application scenarios, and Section~\ref{sec:discussion} discusses the
%main implications of our design.  Finally, Section~\ref{sec:related}
%presents related work and Section~\ref{sec:conclusion} concludes.

\input{efficient-execution/overview}
\input{efficient-execution/prototype}
\input{efficient-execution/evaluation}
\input{efficient-execution/discussion}
