\chapter{Multi-version Software Updates}
\label{chap:safe-updates}

Software updates are an integral part of the software maintenance
process, with new software versions being released on a continuous
basis.  Unfortunately, software updates often result in failures,
which makes many users reluctant to incorporate the latest patches
made available by developers.  As a result, users rely instead on
outdated versions, which despite their relative stability, miss recent
features and bug fixes and may be plagued by security vulnerabilities.

One of the main reasons for which users hesitate to install updates is
that a significant number of them result in failures.  It is only too
easy to find examples of updates that fix a bug or a security
vulnerability only to introduce another problem in 
the code. Our goal is to improve the software update
process in such a way as to encourage users to upgrade to the latest
software version, without sacrificing the stability of the older
version.

We tackle this problem using a simple but effective multi-version
based approach.  Whenever a new update becomes available, instead of
upgrading the software to the new version, we run the new version in
parallel with the old one; by carefully coordinating their executions
and selecting the behaviour of the more reliable version when they
diverge, we create a more secure and dependable multi-version
application.

We implemented this approach in a prototype system called \mx, which
targets crash bugs in Linux applications running on multi-core
processors. \mx allows a new and an old version of an application to
run concurrently, without requiring any
modifications to the application itself or the operating system, nor any
input from the user.
%
% To achieve this goal, \mx combines static and
% dynamic techniques: it uses static analysis at the binary-level to
% construct mappings between the old and the new versions, 
%
The synchronisation of the two versions is performed at the system
call level, using system call interposition and synchronisation.  When
one of the versions crashes, \mx transparently restarts it via a
lightweight checkpointing mechanism and often allows it to survive the bug
by using the code of the other version.

We evaluate \mx by showing that it can successfully survive crashes in
several real applications, specifically several \coreutils utilities and
two popular servers, \lighttpd and \redis.

Software systems are constantly evolving, with new versions and
patches being released on a continuous basis.  Unfortunately, software
updates present a high risk, with many releases introducing new bugs
and security vulnerabilities.

\section{Overview}

\section{Motivating Example}
\label{sec:example}
\input{safe-updates/example}

\section{Challenges}

\section{Mx Prototype System}
\label{sec:mx}
\input{safe-updates/prototype}

\section{Reliability Evaluation}
\label{sec:reliability-evaluation}
\input{safe-updates/evaluation}

%\section{Introduction}
%\label{sec:intro}
%\input{safe-updates/intro}

%\section{Feasibility Study}
%\label{sec:evolution}
%\input{safe-updates/evolution}

%\section{Mx Prototype System}
%\label{sec:implementation}
%\input{safe-updates/prototype}

%\section{Evaluation}
%\label{sec:evaluation}
%\input{safe-updates/evaluation}

%\section{Discussion}
%\label{sec:discussion}
%\input{safe-updates/discussion}

%\section{Conclusion}
%\label{sec:conclusion}
%\input{safe-updates/conclusion}
