\section{Overview}
\label{evolution:overview}

We used the our infrastructure to understand the evolution of \numSystems
popular open-source applications written in C/C++, over a combined period of
\numYears years. The \numSystems evaluated applications are:

\begin{enumerate}

\item[\gnu~\binutils\footnote{\url{http://www.gnu.org/software/binutils/}}]
is a set of utilities for inspecting and modifying object files,
libraries and binary programs.  We selected for analysis the twelve
utilities from the \stt{binutils} folder (\stt{addr2line}, \stt{ar},
\stt{cxxfilt}, \stt{elfedit}, \stt{nm}, \stt{objcopy}, \stt{objdump},
\stt{ranlib}, \stt{readelf}, \stt{size}, \stt{strings} and \stt{strip}),
which are standard user-level programs under many UNIX distributions.

\item[\beanstalkd\footnote{\url{http://kr.github.io/beanstalkd/}}]
is a simple and fast work queue originally designed for reducing the latency of
page views in high-volume web applications.

\item[\git\footnote{\url{http://git-scm.com/}}]
is one the most popular distributed version control systems used by the
open-source developer community.

\item[\lighttpd\footnote{\url{http://www.lighttpd.net/}}]
is a lightweight web server optimized for high performance environments.

\item[\lighttpdtwo\footnote{\url{http://redmine.lighttpdtwo.net/projects/lighttpdtwo2/}}]
is the new major version of the \lighttpd web server developed entirely from
scratch by the same team of developers.

\item[\memcached\footnote{\url{http://memcached.org/}}]
is a general-purpose distributed memory caching system used by several popular
sites such as Craigslist, Digg and Twitter.

\item[\redis\footnote{\url{http://redis.io/}}]
is a popular key-value data store used by many well-known services such as
Twitter, GitHub and StackOverflow.

\item[\vim\footnote{\url{http://www.vim.org/}}] is arguably one of the most
popular text editors.

\item[\zeromq\footnote{\url{http://zeromq.org/}}]
is a high-performance asynchronous messaging middleware library used by a
number of organisations such as Los Alamos Labs, NASA and CERN.

%\item {\bf GNU diffutils} is a collection of four widely-used
%programs: \stt{diff}, \stt{sdiff}, \stt{diff3} and \stt{cmp}, part of
%many popular UNIX distributions.

\end{enumerate}

The \numSystems applications are representative for C/C++ open-source code: GNU
\binutils are user-level utilities, \git is a version control system,
\beanstalkd, \lighttpdtwo, \memcached and \redis are server applications, \vim
is a text editor while \zeromq is a library.  All applications include a
regression test suite.

\begin{table}[t]
\caption{Summary of applications used in our study.
\textit{ELOC} represents the number of executable lines of code and
\textit{TLOC} the number of lines in test files in the last revision
analysed.}
\begin{center}
\begin{tabular}{llrlr}
\toprule
\multicolumn{1}{c}{}     & \multicolumn{2}{c}{\sc Code}& \multicolumn{2}{c}{\sc Tests} \\
\cmidrule(r){2-3}\cmidrule(l){4-5}
\textsc{Application} & \textsc{Language} & \textsc{ELOC} & \textsc{Language} & \textsc{TLOC}          % & \bf Time         
\\ \midrule
\beanstalkd  & C         & \beanstalkdSize & C        & \beanstalkdTsize  % & \beanstalkdTestTime
\\
\binutils    & C         & \binutilsSize  & DejaGnu   & \binutilsTsize    % & \binutilsTestTime 
\\
\git         & C         & \gitSize       & C/shell   & \gitTsize         % & \gitTestTime 
\\
\lighttpd    & C         & \lighttpdSize  & Perl    & \lighttpdTsize    % & \lighttpdtwoTestTime 
\\
\lighttpdtwo    & C         & \lighttpdtwoSize  & Python    & \lighttpdtwoTsize    % & \lighttpdtwoTestTime 
\\
\memcached   & C         & \memcachedSize & C/Perl    & \memcachedTsize   % & \memcachedTestTime 
\\
\redis       & C         & \redisSize     & Tcl       & \redisTsize       % & \redisTestTime    
\\
\vim         & C         & \vimSize       & Vim script       & \vimTsize      % & \vimTestTime   
\\
\zeromq      & C++       & \zeromqSize    & C++       & \zeromqTsize      % & \zeromqTestTime   
\\ \bottomrule
\end{tabular}
\end{center}
\label{tbl:study-systems}
\end{table}

\subsection{Basic characteristics}

Table~\ref{tbl:study-systems} shows some basic characteristics of these
systems: the language in which the code and tests are written, the number of
executable lines of code (ELOC) and the number of lines of test code (TLOC) in
the last revision analysed. To accurately measure the number of ELOC, we
leveraged the information stored by the compiler in \texttt{gcov} graph files,
while to measure the number of TLOC we did a simple line count of the test
files (using \texttt{cloc}, or \texttt{wc~-l} when \texttt{cloc} cannot detect
the file types).

The code size for these applications varies from only \memcachedSize ELOC for
\memcached to \gitSize ELOC for \git.  The test code is written in a variety of
languages and ranges from \lighttpdtwoTsize lines of Python code for
\lighttpdtwo to \gitTsize lines of C and shell code for \git.  The test code is
36\% larger than the application code in the case of \git, approximately as
large as the application code for \memcached, around 40\% of the application
code for \redis and \zeromq, and only around 10\% and 19\% of the application
code for \lighttpdtwo, \vim and \binutils respectively.  Running the test suite
on the last version takes only a few seconds for \binutils, \lighttpdtwo, \vim
and \zeromq, \memcachedTestTime seconds for \memcached, \redisTestTime seconds
for \redis, and 30 minutes for \git, using a four-core Intel Xeon E3-1280
machine with 16 GB of RAM.

The version control system used by majority of these applications is \git.
Four of these projects---\git, \memcached, \redis, and \zeromq ---are hosted on
the \github\footnote{\url{https://github.com/}} online project site.  The other
two---\binutils and \lighttpdtwo---use their own \git hosting. \lighttpd uses
Subversion, but the project also provides a \git mirror on \github. \vim uses
Mercurial.

\subsection{Selection of revisions}

Our goal was to select a comparable number of revisions across applications.
The methodology was to start from the current version at the day of our
experiments, and select an equal number of previous revisions for all systems.
We only counted revisions which modify executable code, tests or both because
this is what our analyses look at. We decided to select 250 such revisions from
each system because some systems had non-trivial dependency issues further back
than this, which prevented us from properly compiling or running them.  We
still had to install the correct dependencies where appropriate, \eg downgrade
\stt{libev} for older versions of \lighttpdtwo and \stt{libevent} for
\memcached.

Note that not all revisions compile, either due to development errors
%(an example of this would be someone forgetting to add a file) 
or portability issues (\eg system header files differing across OS
distributions).
%% Note that we distinguish between this kind of permanent errors (which
%% disallow us to compile all versions earlier than some revision) and
%% more transient compilation errors that affect only some program
%% versions.  
Redis has the largest number of such transient compilation
errors---\redisTransientCompErrs.  The prevailing reasons are missing
\stt{\#include} directives, \eg \stt{unistd.h} for the \stt{sleep} function,
and compiler warnings subsequently treated as errors.  The missing
\stt{\#include} directives most likely slipped past the developers because on
some systems other \stt{libc} headers cause the missing headers to be
indirectly included. The compiler warnings were generated because newer
compiler versions, such as the one that we used, are more pedantic.  Other
reasons include forgotten files and even missing semicolons.

We decided to fix the errors which had likely not been seen at the time a
particular revision was created, for example by adding the compile flag
\stt{-Wno-error} in \binutils so that warnings do not terminate the build
process. In all situations when we could not compile a revision, we rolled over
the changes to the next revisions until we found one where compilation was
successful.  Revisions which do not successfully compile are not counted
towards the 250 limit.

Another important decision concerns the granularity of the revisions
being considered.  Modern decentralised software repositories based on
version control systems such as \git do not have a linear structure
and the development history is a directed acyclic graph rather than a
simple chain.  Different development styles generate different
development histories; for example, \git, \redis and \zeromq exhibit a
large amount of branching and merging while the other three systems
have a mostly linear history.  Our decision was to focus on the main branch,
and treat each merge into it as a single revision. In other words, we
considered each feature branch a single indivisible unit.  Our
motivation for this decision was twofold: first, development branches
are often spawned by individual developers in order to work on a
certain issue and are often ``private'' until they are merged into the
main branch.  As a result, sub-revisions in such branches are often
unusable or even non-compilable, reflecting work-in-progress.  Second,
the main branch is generally the one tracked by most users, therefore
analysing revisions at this level is a good match in terms of
understanding what problems are seen in the field.  This being said,
there are certainly development styles and research questions that would
require tracking additional branches; however, we believe that for our
benchmarks and research questions this level of granularity provides meaningful
answers.

% On a secondary note, we remark that an additional complication with
% this approach is that version control systems do not associate a
% branch name to each revision, so some detective work might be required
% to follow the main development branch.  However, since
% the projects exhibiting a branching structure are hosted on \github, an implicit central
% integrator exists (the project owner) and we considered their history
% to be the official one, essentially always following the first parent
% in a merge.

\begin{table}[t]
\centering
\caption{Revisions used in our study.
  {\em OK}:~code compiles and tests complete successfully,
  {\em TF}:~some tests fail,
  {\em TO}:~tests time out,
  {\em CF}:~compilation fails,
  {\em Time}:~the number of months analysed.}
\begin{tabular}{lrrrrr}
\toprule
\multicolumn{1}{c}{}          &       \multicolumn{3}{c}{\sc OK+TF+TO=250}                 &            \multicolumn{2}{c}{}                   \\
\cmidrule{2-4}
\textsc{Application} & \textsc{OK} & \textsc{TF} & \textsc{TO} & \textsc{CF} & \textsc{Time}           \\
\midrule
\beanstalkd  &  \beanstalkdOK & \beanstalkdTransientTestErrs & \beanstalkdTransientTestTimeouts & \beanstalkdTransientCompErrs  &  {\beanstalkdTimespan}mo \\
\binutils    &  \binutilsOK   & \binutilsTransientTestErrs  & \binutilsTransientTestTimeouts  & \binutilsTransientCompErrs  &  {\binutilsTimespan}mo \\
\git         &  \gitOK        & \gitTransientTestErrs       & \gitTransientTestTimeouts       & \gitTransientCompErrs       &  {\gitTimespan}mo  \\
\lighttpd    &  \lighttpdOK   & \lighttpdTransientTestErrs  & \lighttpdTransientTestTimeouts  & \lighttpdTransientCompErrs  &  {\lighttpdTimespan}mo  \\
\lighttpdtwo    &  \lighttpdtwoOK   & \lighttpdtwoTransientTestErrs  & \lighttpdtwoTransientTestTimeouts  & \lighttpdtwoTransientCompErrs  &  {\lighttpdtwoTimespan}mo  \\
\memcached   &  \memcachedOK  & \memcachedTransientTestErrs & \memcachedTransientTestTimeouts & \memcachedTransientCompErrs &  {\memcachedTimespan}mo \\
\redis       &  \redisOK      & \redisTransientTestErrs     & \redisTransientTestTimeouts     & \redisTransientCompErrs     &  {\redisTimespan}mo     \\
\vim         &  \vimOK        & \vimTransientTestErrs       & \vimTransientTestTimeouts       & \vimTransientCompErrs    &  {\vimTimespan}mo    \\
\zeromq      &  \zeromqOK     & \zeromqTransientTestErrs    & \zeromqTransientTestTimeouts    & \zeromqTransientCompErrs    &  {\zeromqTimespan}mo    \\
\bottomrule
\end{tabular}
\label{tbl:revisions}
\end{table}


Table~\ref{tbl:revisions} summarises the revisions that we selected:
they are grouped into those that compile and pass all the tests (\emph{OK}),
compile but fail some tests (\emph{TF}), compile but time out while running the
test suite (\emph{TO}), and fail to compile (\emph{CF}).  The time limit that
we enforced was empirically selected for each system such that it is large
enough to allow a correct revision to complete all tests. As shown in the
table, timeouts were a rare occurrence, with at most one occurrence per
application.

Table~\ref{tbl:revisions} also shows the development time span
considered, which ranges from only 5-6 months for \git and \redis,
which had a fast-paced development during this period, to almost 4
years for \memcached. The age of the projects at the first version
that we analysed ranges from a little over 2 years for \lighttpdtwo,
to 11 years for \binutils.
%6 years memcached, 4 years redis, 2yr 9mo zeromq, 8 years git

\subsection{Revision setup}

All the programs analysed were compiled to record coverage information. In
addition, we disabled compiler optimisations, which generally interact poorly
with coverage measurements. We used existing build targets and configuration
options if available, otherwise we configured the application with the flags
\lstinline`CFLAGS='-O0 -coverage'` and \lstinline`LDFLAGS=-coverage`. All code
from the system headers, \ie \stt{/usr/include/} was excluded from the results.

Each revision was run in a virtualised environment based on 64-bit version of
Ubuntu 12.10 (12.04.3 for \git) running inside an \lxc container.  To take
advantage of the inherent parallelism of this approach, the containers were
spawned in one of 28 long-running \xen VMs, each with a 4~Ghz CPU, 6~GB of RAM,
and 20~GB of storage, running a 64-bit version of Ubuntu 12.04.3.

% The following sections present the main findings of our analysis.  They first
% reiterate and then examine in detail our target research questions (RQs).
